---
title: "ML Final Project - Medical Student Mental Health and Burnout Levels"
author: "Group 3: Haoran Lu, Emma Krolicki"
date: "2024-04-15"
output: pdf_document
---

# Loading required packages
```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(caret)
library(ISLR2)
library(randomForest)
library(ltm)
library(scales)
library(ROSE)
library(smotefamily)
library(pROC)
library(gbm)
```

# Reading and recoding the dataset
```{r}
# Reading dataset and notebook
data <- read.csv("Data Carrard et al. 2022 MedTeach.csv")
codebook <- read.csv2("Codebook Carrard et al. 2022 MedTeach.csv")
# id (Any duplicates)
anyDuplicated(data$id) # No
# age (Should be normal age)
summary(data$age) # 17 to 49, OK
# year (Should be 1 to 6)
summary(data$year) # True
# sex (Should be 1, 2, or 3)
table(data$sex) # True
# glang (Should be one of the languages)
table(data$glang) # True, but too many languages for data analysis
# part (Should be 0 or 1)
table(data$part) # True
# job (Should be 0 or 1)
table(data$job) # True
# stud_h (Hours per week, should be considered "normal")
summary(data$stud_h) # 0 to 70. Max = 70, could be a hard working student
# health (Should be 1 to 5)
table(data$health) # True
# psyt (Should be 0 or 1)
table(data$psyt) # True
# jspe (Should be 20 to 140)
summary(data$jspe) # True
# qcae_cog (Should be 19 to 76)
summary(data$qcae_cog) # True
# qcae_aff (Should be 12 to 48)
summary(data$qcae_aff) # True
# amsp (Should be 5 to 35)
summary(data$amsp) # True
# erec_mean (Should be 0 to 1)
summary(data$erec_mean) # True
# cesd (Should be 0 to 60)
summary(data$cesd) # True
# stai_t (Should be 20 to 80)
summary(data$stai_t) # True
# mbi_ex (Should be 0 to 30)
summary(data$mbi_ex) # True
# mbi_cy (Should be 0 to 24)
summary(data$mbi_cy) # True
# mbi_ea (Should be 0 to 36)
summary(data$mbi_ea) # True

# Specifying columns needed recoding
c <- c(3,4,5,6,7,9,10)

# Recoding funtion
apply_recoding <- function(data, codebook) {
  for (i in c) {
    variable <- codebook$Variable.Name[i]
    scale_info <- codebook$Variable.Scale[i]
    
    parts <- strsplit(scale_info, "; ")[[1]]
    recode_list <- setNames(
    lapply(parts, function(x) strsplit(x, "=")[[1]][2]),
    lapply(parts, function(x) strsplit(x, "=")[[1]][1])
      )
    recode_list <- unlist(recode_list)
    data[[variable]] <- recode(data[[variable]], !!!recode_list)
    }
   return(data)
}

# Recoding dataset
data <- apply_recoding(data, codebook)
```

### Define burnout level as dichotomous variable
# Burnout level was defined by both high emotional exhaustion (>26) and high depersonalization (>12)
# Williamson, K., Lank, P. M., Cheema, N., Hartman, N., Lovell, E. O., & Emergency Medicine Education Research Alliance (EMERA) (2018). Comparing the Maslach Burnout Inventory to Other Well-Being Instruments in Emergency Medicine Residents. Journal of graduate medical education, 10(5), 532–536.

# Reverse recoding for mbi_ea, add three tests up as a new continuous variable indicating burnout
# https://www.sciencedirect.com/science/article/pii/S0738399121003918#sec0010
```{r}
# Reverse mbi_ea
data$mbi_ea_re <- (36 - data$mbi_ea)
data$mbi_total <- (data$mbi_cy + data$mbi_ea_re + data$mbi_ex)

# Binary classification
data$burnout <- ifelse(data$mbi_ex >= 26 & data$mbi_cy >= 12, "Yes", "No")
data$burnout <- as.factor(data$burnout)

# Recoding of `glang` less than 10 as `Other`
data$glang <- ifelse(table(data$glang)[data$glang] > 10, 
                            as.character(data$glang), "Other")
table(data$glang)%>%sort(decreasing = T)

data$sex <- factor(data$sex)
data$health <- factor(data$health)
data$year <- factor(data$year)
data$part <- factor(data$part)
data$job <- factor(data$job)
data$psyt <- factor(data$psyt)
data$glang <- factor(data$glang)
data$burnout <- factor(data$burnout, levels = c("No", "Yes"))
```

# Descriptive analysis of the dataset
```{r}
# Distribution of values in column with categorical data
cat_var <- c(3,4,5,6,7,9,10,23)
cat_col_names <- names(data)[cat_var]

# Plotting each categorical column with percentages
for (col in cat_col_names) {
  ggplot(data = data, aes_string(x = col, fill = col)) + 
    geom_bar() + 
    geom_text(aes(label = scales::percent(..count../sum(..count..))), 
              position = position_fill(vjust = 0.5), stat = "count") +
    scale_y_continuous(labels = scales::percent_format()) +
    theme_minimal() +
    labs(title = paste("Percentage Bar Chart of", "'", col, "'"),
         x = col, y = "Percentage") +
    theme(plot.title = element_text(hjust = 0.5)) -> p
  print(p)
}

# Correlation between `mbi_total` and other numeric variables
library(dplyr)
library(corrplot)
data_num <- data %>% select_if(is.numeric)
data_num <- data_num[,-c(1,11,12,13,14)]
cor_matrix <- cor(data_num)
col <- colorRampPalette(c("white", "red"))(200)
corrplot(cor_matrix, col = col, 
         method = "color", type = "upper", order = "hclust", 
         tl.col = "black", tl.cex = 0.8, tl.srt = 45, 
         addCoef.col = "black")

# Distribution of mbi_total
hist(data$mbi_total, 
     main = "Histogram of `mbi_total`",
     xlab = "`mbi_total`",
     freq = FALSE,
     breaks = 20,
     col = "lightgray")
curve(dnorm(x, mean = mean(data$mbi_total), 
            sd = sd(data$mbi_total)), 
      add = TRUE, col = "red", lwd = 2)
```


# Logistic regression
```{r}
set.seed(1)
# Cross validation
train_control <- trainControl(
  method = "cv",  
  number = 5,  
  classProbs = TRUE,  
  summaryFunction = twoClassSummary,  
  savePredictions = "final",  
  sampling = "up"  
)

# Fit the model
model_cv <- train(
  burnout ~ age + year + sex + glang + part + job + stud_h + health + psyt + 
    jspe + qcae_cog + qcae_aff + amsp + erec_mean + cesd + stai_t,
  data = data,
  method = "glm",
  family = "binomial",
  trControl = train_control,
  metric = "ROC"
)

# View model performance
summary(model_cv)
predictions <- predict(model_cv, newdata = data, type = "prob")
prob_yes <- predictions[,2]
roc_final <- roc(data$burnout, prob_yes)
auc_lg <- auc(roc_final)
ggroc(roc_final, size = 2, alpha = 0.5) +
  ggtitle("ROC Curve for Predicting 'Burnout'") + 
  geom_label(aes(x = 0.2, y = 0.2, label = sprintf("AUC = %.2f", auc_lg)), 
                       fill = "white", 
                       size = 6,       
                       color = "black" 
                      )

final_model <- model_cv$finalModel
pre_d <- predict(model_cv, newdata = data)
conf_matrix <- confusionMatrix(pre_d, data$burnout, positive = "Yes")
summary(final_model)

# Oversampling
data_balanced <- ovun.sample(burnout ~ ., 
                             data = data, method = "over", 
                             N = 1686, seed = 1)$data
model_cv_balanced <- train(
  burnout ~ age + year + sex + glang + part + job + stud_h + health + psyt + 
    jspe + qcae_cog + qcae_aff + amsp + erec_mean + cesd + stai_t,
  data = data_balanced,
  method = "glm",
  family = "binomial",
  trControl = train_control,
  metric = "ROC"
)

predictions_b <- predict(model_cv_balanced, newdata = data_balanced, 
                         type = "prob")
prob_yes_b <- predictions_b[,2]
roc_final_b <- roc(data_balanced$burnout, prob_yes_b)
auc_lg_b <- auc(roc_final_b)
ggroc(roc_final_b, size = 2, alpha = 0.5) +
  ggtitle("ROC Curve for Predicting 'Burnout'") + 
  geom_label(aes(x = 0.2, y = 0.2, label = sprintf("AUC = %.2f", auc_lg_b)), 
                       fill = "white", 
                       size = 6,       
                       color = "black" 
                      )
pre_d_b <- predict(model_cv_balanced, newdata = data_balanced)
conf_matrix_b <- confusionMatrix(pre_d_b, data_balanced$burnout, 
                               positive = "Yes")
final_model_b <- model_cv_balanced$finalModel
summary(final_model_b)
```

# Random forest models
```{r}
# Create training and testing sets
split <- as.vector(createDataPartition(data$burnout, p = 0.8, list = FALSE))
train_set <- data[split, ]
test_set <- data[-split, ]

# Randomforest model using all variables
rf_1 <- randomForest(burnout ~ age + year + sex + glang + part + job + 
                       stud_h + health + psyt + jspe + qcae_cog + qcae_aff +
                       amsp + erec_mean + cesd + stai_t, 
                     data = train_set, 
                     ntree = 50, importance = TRUE)
rf_1
mean(predict(rf_1, newdata = train_set) == train_set$burnout)
mean(predict(rf_1, newdata = test_set) == test_set$burnout)
confusionMatrix(predict(rf_1, newdata = train_set), train_set$burnout)
confusionMatrix(predict(rf_1, newdata = test_set), test_set$burnout)

# Using balanced training set
balan_rf <- ovun.sample(burnout ~ ., data = train_set, method = "over", 
                        N = 1350, seed = 1)$data
rf_2 <- randomForest(burnout ~ age + year + sex + glang + part + job + stud_h + 
                       health + psyt + jspe + qcae_cog + qcae_aff + amsp + 
                       erec_mean + cesd + stai_t, 
                     data = balan_rf, 
                     ntree = 50, importance = TRUE)
rf_2
plot(rf_2,
     main = "Error Rate vs Number of Trees")
mean(predict(rf_2, newdata = train_set) == train_set$burnout)
mean(predict(rf_2, newdata = test_set) == test_set$burnout)
confusionMatrix(predict(rf_2, newdata = test_set), test_set$burnout)
```

# Predicting burnout level as numeric variable
```{r}
set.seed(1)
# Use randomForest to predict burnout level as numeric variable, balanced dataset
rf_num <- randomForest(mbi_total ~ age + year + sex + glang + part + job + 
                       stud_h + health + psyt + jspe + qcae_cog + qcae_aff +
                       amsp + erec_mean + cesd + stai_t, 
                       data = train_set, 
                       ntree = 50, importance = TRUE)
rf_num
pred_rfn <- predict(rf_num, newdata = test_set)
pred_rft <- predict(rf_num, newdata = train_set)
MSE_test <- mean((pred_rfn - test_set$mbi_total)^2)
mean((pred_rft - train_set$mbi_total)^2)

# Calculate R^2 and RMSE using postResample
R2_test <- postResample(pred = pred_rfn, obs = test_set$mbi_total)
R2_train <- postResample(pred = pred_rft, obs = train_set$mbi_total)

SSE <- sum((pred_rfn - test_set$mbi_total)^2)  
SST <- sum((test_set$mbi_total - mean(test_set$mbi_total))^2)  
R_squared <- 1 - SSE/SST
n <- nrow(test_set)
p <- length(rf_num$coefficients) - 1

test_adjr2 <- 1 - ((1 - R_squared) * (n - 1) / (n - p - 1))

# Unbalanced dataset
rf_num_un <- randomForest(mbi_total ~ age + year + sex + glang + part + job + 
                       stud_h + health + psyt + jspe + qcae_cog + qcae_aff +
                       amsp + erec_mean + cesd + stai_t, 
                       data = train_set, 
                       ntree = 50, importance = TRUE)

mean((predict(rf_num_un, newdata = train_set) - train_set$mbi_total)^2)
mean((predict(rf_num_un, newdata = test_set) - test_set$mbi_total)^2)
pred_rfn2 <- predict(rf_num_un, newdata = test_set)

# Calculate R^2 and RMSE using postResample
results2 <- postResample(pred = pred_rfn2, obs = test_set$mbi_total)

SSE <- sum((pred_rfn2 - test_set$mbi_total)^2)  
SST <- sum((test_set$mbi_total - mean(test_set$mbi_total))^2)  
R_squared <- 1 - SSE/SST
n <- nrow(test_set)
p <- length(rf_num_un$coefficients) - 1

test_adjr22 <- 1 - ((1 - R_squared) * (n - 1) / (n - p - 1))

# Gradient boosting 
train_set[] <- lapply(train_set, 
                      function(x) if(is.character(x)) factor(x) else x)
test_set[] <- lapply(test_set, 
                     function(x) if(is.character(x)) factor(x) else x)
gbm_1 <- gbm(mbi_total ~ age + year + sex + glang + part + job + 
                       stud_h + health + psyt + jspe + qcae_cog + qcae_aff +
                       amsp + erec_mean + cesd + stai_t,
             data = train_set,  
             distribution = "gaussian",  
             n.trees = 500,  
             interaction.depth = 1,  
             cv.folds = 5)

gbm1_test <- predict(gbm_1, newdata = test_set, n.trees = 500)
r2_po_test <- postResample(pred = gbm1_test, obs = test_set$mbi_total)
MSE_boost <- mean((gbm1_test - test_set$mbi_total)^2)
```
























